---
title: "Installing the ktest package"
author: "Ghislain DURIF"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installing the ktest package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Package requirements

- R version 4+
- Python version 3+

**Important:** Python is a requirement as an intern machinery for the package to work but you will not need to create nor manipulate Python codes to use the `ktest` R package.

> **Note:** if you don't have Python on your system: when configuring the `ktest` R package (c.f. [below](#first-time-configuration)), the `reticulate` R package will offer to install Python on your system.

The `ktest` R package is using the `ktest` Python package under the hood thanks to the [`reticulate`](https://CRAN.R-project.org/package=reticulate) R package that provides an "R Interface to Python".

## Installation

You can install the development version of `ktest` with the following commands:

```{r setup-install-dev, eval=FALSE}
install.packages("remotes")
remotes::install_github("LMJL-Alea/ktest", ref = "main", subdir = "r-pkg")
```

> **Note:** `ktest` is not available on CRAN at the moment but will be shortly.

## First-time configuration

After installing the `ktest` R package, you need to run the following commands (**once**) to complete the setup (and install the `ktest` Python package on your system):

> :warning: To avoid messing with your Python system or user environment, we recommend to use a dedicated Python environment for `ktest` (c.f. [next section](#using-a-python-environment)) :warning:

```{r setup-ktest-simple, eval=FALSE}
# load ktest R package
library(ktest)
# install ktest package python requirements
install_ktest()
# check ktest configuration
check_ktest()
```

### Using a Python environment

Here are the commands to be run (once) to configure the `ktest` package using a dedicated Python **virtual environment**:

```{r setup-ktest-venv, eval=FALSE}
# load ktest R package
library(ktest)
# create dedicated Python virtual environment
reticulate::virtualenv_create("ktest")
# activate the python environment
reticulate::use_virtualenv(virtualenv = "ktest", required = TRUE)
# verify python version
reticulate::py_config()
# install ktest package python requirements
install_ktest(method = "virtualenv", envname = "ktest")
# check ktest configuration
check_ktest()
```

> **Note:** if you are a Miniconda/Anaconda Python distribution user, you can either user a Python virtual environment (c.f. above) or a **Conda environment** with the same results. Please refer to the `reticulate` [documentation](https://rstudio.github.io/reticulate/articles/versions.html#providing-hints) in this case.


### Managing Python

To check which **version of Python** you are using through `reticulate`, you can run:

```{r py-config, eval=FALSE}
reticulate::py_discover_config()
```

> **Note:** To get more information about **managing** which **version of Python** you are using, you can refer to `reticulate` [documentation](https://rstudio.github.io/reticulate/articles/versions.html) about "Python Version Configuration".

## Usage

Once the `ktest` package is installed and configured, to use it, you only need to load it like any other R package:
```{r load-ktest, eval=FALSE}
library(ktest)
kt <- ktest(...) # see other vignettes
```

:warning: If you are using a **dedicated Python environment**, you also need to load it every time before using `ktest`:
```{r load-ktest-env, eval=FALSE}
library(ktest)
reticulate::use_virtualenv(virtualenv = "ktest", required = TRUE)
kt <- ktest(...) # see other vignettes
```
