---
title: "Installing the ktest package"
author: "Ghislain DURIF"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installing the ktest package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Requirements

- R version 4+
- Python version 3+ 

> **Note:** if you don't have Python on your system: when configuring the `ktest` R package (c.f. [below](#additional-setup)), the `reticulate` R package will offer to install Python on your system.

The `ktest` R package is using the `ktest` python package under the hood thanks to the [`reticulate`](https://CRAN.R-project.org/package=reticulate) R package that provides an "R Interface to Python".

### Install `ktest` R package

You can install the development version of `ktest` with the following command:

```{r setup-install-dev, eval=FALSE}
remotes::install_github("AnthoOzier/ktest", ref = "rktest_dev", subdir = "r-pkg")
```

> **Note:** `ktest` is not available on CRAN at the moment but will be shortly.

### Additional setup

After installing the `ktest` R package, you need to run the following command (**once**) to complete the setup (and install the `ktest` Python package on your system):

> :warning: to avoid messing with your Python system environment, we recommend to use a dedicated Python virtual environment (c.f. [next section](#using-python-virtual-environment)) :warning:

```{r setup-ktest-simple, eval=FALSE}
library(ktest)
install_pyktest()
```

### Managing Python

If you need more **details** about which **version of Python** you are using through `reticulate`, you can run:

```{r py-config, eval=FALSE}
reticulate::py_config()
reticulate::py_discover_config()
```

> **Note:** To get more information about **managing** which **version of Python** you are using, you can refer to `reticulate` [documentation](https://rstudio.github.io/reticulate/articles/versions.html) about "Python Version Configuration".

#### Using Python virtual environment

1. To create a dedicated Python environment, run the following command (once):
```{r setup-python-env, eval=FALSE}
# create dedicated virtual environment
reticulate::virtualenv_create("ktest")
```

2. To load the Python environment, you should run the following command:
```{r load-python-env, eval=FALSE}
# activate python environment
reticulate::use_virtualenv(virtualenv = "ktest", required = TRUE)
```

3. Setup `ktest` Python dependency inside the dedicated environment (to be run once):
```{r setup-ktest, eval=FALSE}
library(ktest)
install_pyktest(method = "virtualenv", envname = "ktest")
```

> **Note:** you can also use a Conda environment if you use a Miniconda/Anaconda Python distribution, c.f. `reticulate` [documentation](https://rstudio.github.io/reticulate/articles/versions.html) about "Python Version Configuration".

### Usage

Once the `ktest` package is installed and configured, to use it, you only need to load it like any other R package:
```{r load-ktest, eval=FALSE}
library(ktest)
kt <- ktest(...) # see other vignettes
```

If you use a dedicated Python environment, you also need to load it before using `ktest`:
```{r load-ktest-env, eval=FALSE}
reticulate::use_virtualenv(virtualenv = "ktest", required = TRUE)
library(ktest)
kt <- ktest(...) # see other vignettes
```
