---
title: "Multivariate testing"
author: "Ghislain DURIF"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multivariate testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, eval=FALSE}
library(ktest)
```

```{r, include=FALSE}
proj_dir <- rprojroot::find_root(".git/index")
pkg_dir <- file.path(proj_dir, "r-pkg")
devtools::load_all(pkg_dir)
```

## Load data

Load data and metadata:

```{r multi-data}
# cell expression data
sc_df <- read.table("data/data2.csv", row.names = 1, sep = ",", header = TRUE)
# cell metadata
meta_sc_df <- read.table("data/metadata2.csv", row.names = 1, sep = ",", header = TRUE)
```

## Ktest definition 

Initialize the `ktest` object:

- `data` is the dataframe of data.
- `metadata` is the dataframe of metadata.
- `condition` is the column of `metadata` containing the labels to test. 
- `samples` is the couple of samples to test in the column `condition` of `metadata`
- Set `nystrom` to `TRUE` to use the Nystrom approximation. 
- Set `verbose >0` to follow the steps of the initialization of the `ktest` object.  

```{r multi-init}
kt <- ktest(
    sc_df, meta_sc_df,
    condition='condition', samples=c('0H','48HREV'), verbose=1
)
```

## Multivariate testing 

Perform a multivariate test on the data:
```{r multi-test}
# R call
multivariate_test(kt, verbose=1)
# equivalent Python call through reticulate
kt$multivariate_test(verbose=1)
```

### Print test results 

- `long` : if `TRUE`, the results are printed for several truncations
- `t` : truncation associated to the printed p-value if `long` is `FALSE`.
- `ts` : list of truncations associated to the printed p-values if `long` is `TRUE`.

```{r multi-result}
# R call
print_multivariate_test_results(kt, long = TRUE, ts = c(1,2,3))
# equivalent Python call through reticulate
kt$print_multivariate_test_results(long = TRUE, ts = c(1,2,3))
```


### Get p-values table with respect to the truncation parameter. 

- `contrib`: if true, returns the p-value associated to each principal component individually. Returns the p-value associated to the `kfda` statistic otherwise. 
- `log`: if `TRUE`, returns the log p-values. 

```{r multi-pval}
# R call
pval <- get_pvalue(kt, contrib = TRUE, log = FALSE)
# equivalent Python call through reticulate
pval <- kt$get_pvalue(contrib = TRUE, log = FALSE)

head(pval)
```

### Plot p-values with respect to truncation

FIX ME!

### Other tests

The default p-value is the asymptotic p-value of the `kfda` statistic associated to the truncation parameter `t=10`. To change the default truncation parameter, use the class function `set_truncation`. 


```{r multi-change-trunc}
# Python call through reticulate
kt$set_truncation(5L)
kt$multivariate_test(verbose=1)
```

The asymptotic p-value is available for the `kfda` statistic only. 
The permutation p-value is available for the `kfda` statistic and the `mmd` statistic.

Compute the permutation p-value associated to the chosen statistic:

- `stat` : chosen statistic among (`kfda`,`mmd`)
- `permutation`: if True, compute the permutation p-value (automatically True if `stat` is `mmd`)
- `n_permutations` : set the number of permutations.
- `seed_permutation` : define the number of random seeds. 
- `n_jobs_permutation` : number of CPU to use for parallelized computation

```{r multi-mmd}
# Python call through reticulate
kt$multivariate_test(
    stat='mmd',
    permutation=TRUE,
    n_permutations=2000L,
    seed_permutation=123L,
    n_jobs_permutation=3L,
    verbose=1L)
```

Results:
```{r multi-mmd-result}
# Python call through reticulate
kt$stat
kt$get_pvalue_name()
kt$dict_pval_mmd[kt$get_pvalue_name()]
kt$get_pvalue()
```

### Nystrom approximation

Use the Nystrom approximation to reduce the computational cost:
```{r multi-nystrom}
# Python call through reticulate
kt$set_test_params(nystrom=TRUE)
kt$multivariate_test(verbose=1)
```

Tune each parameter of the Nystrom approximation with the function `set_test_params()`:

- `nystrom` : if True, use the Nystrom approximation
- `n_landmarks` : number of landmarks to use (subsample used to compute the anchors).
- `n_anchors` : number of anchors to use (eigenvectors associated to a matrix defined with the landmarks).
- `landmark_method` : how to choose the landmarks among (`random`,`kmeans`)
- `anchor_basis` : matrix used to define the anchors among 
  + `k` : second moment of the landmarks.
  + `s` : total covariance of the landmarks.
  + `w` : within group covariance of the landmarks.


```{r multi-tune-nystrom}
# Python call through reticulate
kt$set_test_params(
    nystrom=TRUE,
    n_landmarks=50L,
    n_anchors=50L,
    landmark_method='random',
    anchor_basis='s'
)
kt$multivariate_test(verbose=1)
```

### Kernel function choice

**TODO**

The default kernel is the RBF kernel with median bandwidth. Use `init_kernel_params()` to specify the kernel function with a `dict` of specifications:

- `function` is either a string among [`gauss`,`linear`] or a user specified kernel function.
  + if `function` is `gauss`:
    - `bandwidth` is either the string `median` or a float. 
    - if `bandwidth` is `median`: 
      + `median_coef` is the coefficient such that `bandwidth=median_coef x median` (default is `1`)
- `kernel_name` is the name of the kernel used.  


```python
import torch
from pyktest.base import init_kernel_params

# gauss kernel with median/2 bandwidth
kernel1 = init_kernel_params(function = 'gauss', bandwidth = 'median', median_coef = 1/2)

# linear kernel 
kernel2 = init_kernel_params(function='linear')

# user specified kernel
kernel3 = init_kernel_params(function=lambda x,y: torch.cdist(x,y).exp()+1,kernel_name='useless_kernel')
```


```python
kt = Ktest(data,
           metadata,
           condition='condition',
           samples=['0H','48HREV'],
#            kernel=kernel3,
           verbose=1)

kt.multivariate_test(verbose=1)
```

    - Add data 'data' to Ktest, dimensions=(685, 83)
    - Set test data info (0H,48HREV from condition)
    - Define kernel function (gauss)
    - Initialize kfdat
    	cov : standard 
    	mmd : standard
    - Diagonalize within covariance centered gram
    - Compute within covariance centered gram
    - Compute kfda statistic
    
    ___Multivariate kfda test results___
    Asymptotic p-value(truncation) for multivariate testing : 
    	p-value(10) = 3.9e-23 (kfda=130.41)

