---
title: "Development actions history"
author: "Ghislain DURIF"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Usethis setup

```{r dev-setup}
path <- rprojroot::find_root(".git/index")
usethis::proj_set(path)
```

## Package development

### Initialization (to be run once)

```{r dev-init}
# create package template
usethis::create_package("rktest", rstudio = FALSE, open = FALSE)
# use Rmd for README generation
usethis::use_readme_rmd()
```

```{r dev-git}

### Metadata

```{r dev-metadata}
# description
use_description(
    fields = list(
        Title = "Kernal based statistical testing",
        Description = "Write Me",
        `Authors@R` = c(
            person(
                "Anthony", "Ozier-Lafontaine", 
                email = "anthony.ozier-lafontaine@ec-nantes.fr", 
                role = c("aut")),
            person(
                "Ghislain", "Durif", email = "gd.dev@libertymail.net", 
                role = c("aut", "cre"), 
                comment = c(ORCID = "0000-0003-2567-1401"))
        ),
        URL = stringr::str_c(
            "https://plmlab.math.cnrs.fr/gdurif/funStatTest/", 
            "https://gdurif.pages.math.cnrs.fr/funStatTest/", 
            sep = ","
        ),
        Version = "1.0.0",
        "Date/Publication" = format(Sys.time(), "%Y/%m/%d")
    ),
    check_name = TRUE, 
    roxygen = TRUE
)
# License
usethis::use_mit_license()
# Roxygen markdown
usethis::use_roxygen_md()
```