---
title: "Univariate testing"
author: "Ghislain DURIF"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r python}
library(reticulate)
# create virtualenv
virtualenv_create("ktest")
# activate python environment
use_virtualenv(virtualenv = "ktest", required = TRUE)

# install Python package
virtualenv_install(
    envname = "ktest",
    packages = "pyktest @ git+https://github.com/AnthoOzier/ktest@rktest_dev#subdirectory=pyktest"
)
```

Import

```python
import pandas as pd
from ktest.tester import Ktest
```

```{r import}
pyktest <- reticulate::import("pyktest")
```

### Perform univariate testing (differential expression analysis) on all the genes


```python
data = pd.read_csv('data.csv',index_col=0)
metadata = pd.read_csv('metadata.csv',index_col=0)
kt = Ktest(data,metadata,condition='condition',nystrom=True)
```

Perform univariate testing on all the genes.


- `n_jobs` : Parallelize the jobs on `n_jobs` CPUs if possible
- `save_path` : Save the results in the `save_path` directory 
- `lots` : Divide the total set of variables to test in sets of size `lots` to save intermediate results.
- `name` : Use `name` to refer to this set of tests if you aim at testing different groups of variables separately (See section **Perform multiple testing on different pathways** for more details).  
- `truncations_of_interest` : the test results to keep (default =10 mais pour l'instant je vais garder défaut = 1:30) 
- `diagnostics` : Set as `True` to compute the diagnostic metrics for each variables 
- `t_diagnostics` : The diagnostic metrics are saved for truncations 1 to `t_diagnostic`  
- `verbose` : Whatch the job in progress with `verbose > 0`



```python
kt.univariate_test(n_jobs=4,lots=20,verbose=1,name='all_variables',
                  save_path='./')
```

    - Load univariate test results from 
    	dir: ./
    	file: all_variables_ny_lmrandom_m33_basisw_datacondition_univariate.csv
    - Loaded univariate test results : (83, 25)
    - Update var with 25 columns
    - No variable to test
    - Load univariate test results from 
    	dir: ./
    	file: all_variables_ny_lmrandom_m33_basisw_datacondition_univariate.csv
    - Loaded univariate test results : (83, 25)
    - Saving univariate test results of all_variables in 
    	dir: ./
    	file:all_variables_ny_lmrandom_m33_basisw_datacondition_univariate.csv ((83, 25))


The function does not repeat the computation if not necessary 


```python
kt.univariate_test(verbose=1)
```

    - No variable to test


---

Print summarized results :
- `t` is the truncation parameter of interest (default is 10).
- Set `long` as `True` for detailled information.
- set `corrected` as `False` to have the raw p-values not corrected with the Benjamini-Hochberg approach for multiple testing.
- `ntop` number of top DE variables to print.
- `threshold` : test rejection threshold. A variable is considered as DE if `p-value < threshold`.
- Set `log2fc` as `True` to print the log2fc of the displayed DE variables if possible. 


```python
kt.print_univariate_test_results(ntop=3)
```

    
    ___Univariate tests results___
    5 DE genes for t=10 and threshold=0.05 (with BH correction)
    Top 3 DE genes (with BH correction): 
    PLS1, SLC6A9, STX12


Get the resulting p-values in a Pandas.Series :
- `t` is the truncation parameter of interest
- set `corrected` as `False` to have the raw p-values not corrected with the Benjamini-Hochberg approach for multiple testing.


```python
kt.get_pvals_univariate(verbose=2)
```

    Warning : only 21 variables were corrected out of 83 tested variables





    PLS1       2.841728e-07
    SLC6A9     2.448390e-06
    STX12      2.448390e-06
    NCOA4      1.007764e-05
    PIK3CG     1.633637e-02
                   ...     
    TPP1                NaN
    TTYH2               NaN
    UCK1                NaN
    VDAC3               NaN
    XPNPEP1             NaN
    Name: all_variables_ny_lmrandom_m33_basisw_datacondition_t10_pvalBHc, Length: 83, dtype: float64



Get the resulting p-values of DE genes only :
- `t` is the truncation parameter of interest.
- set `corrected` as `False` to have the raw p-values not corrected with the Benjamini-Hochberg approach for multiple testing.
- `threshold` : test rejection threshold. A variable is considered as DE if `p-value < threshold`.


```python
kt.get_DE_genes(verbose=1)
```

    Warning : only 21 variables were corrected out of 83 tested variables





    PLS1      2.841728e-07
    SLC6A9    2.448390e-06
    STX12     2.448390e-06
    NCOA4     1.007764e-05
    PIK3CG    1.633637e-02
    Name: all_variables_ny_lmrandom_m33_basisw_datacondition_t10_pvalBHc, dtype: float64



###  Perform multiple testing on different pathways

Define the pathways of interest (the following pathways are random, they do not have biological meaning). 


```python
pathway1 = ['CTCF', 'CTSA', 'CYP51A1', 'DCP1A', 'DCTD', 'DHCR24', 'DHCR7', 'DPP7']
pathway2 = ['MAPK12', 'MFSD2B', 'MID2', 'MKNK2', 'MTFR1', 'MVD', 'MYO1G', 'NCOA4']
pathway3 = ['NSDHL', 'PDLIM7', 'PIK3CG', 'PLAG1', 'PLS1', 'PLS3', 'PPP1R15B']

```

Perform univariate testing on each pathway 


```python
kt.univariate_test(variables_to_test=pathway1,name='pathway1')
kt.univariate_test(variables_to_test=pathway2,name='pathway2')
kt.univariate_test(variables_to_test=pathway3,name='pathway3',save_path='./',verbose=1)
```

    
    
    - Load univariate test results from 
    	dir: ./
    	file: pathway3_ny_lmrandom_m33_basisw_datacondition_univariate.csv
    - Loaded univariate test results : (7, 21)
    - Update var with 21 columns
    - No variable to test
    - Load univariate test results from 
    	dir: ./
    	file: pathway3_ny_lmrandom_m33_basisw_datacondition_univariate.csv
    - Loaded univariate test results : (7, 21)
    - Saving univariate test results of pathway3 in 
    	dir: ./
    	file:pathway3_ny_lmrandom_m33_basisw_datacondition_univariate.csv ((7, 21))


By default, the functions `print_univariate_test_results`, `get_pvals_univariate` and `get_DEgenes` refer to the last tested set of variables (here it is `pathway3`): 


```python
kt.print_univariate_test_results()
```

    
    ___Univariate tests results___
    7 tested variables out of 83 variables
    0 DE genes for t=10 and threshold=0.05 (with BH correction)


Print or get the results of `pathway1` using parameter `name` 


```python
kt.print_univariate_test_results(name='pathway1')
kt.get_pvals_univariate(t=2,name='pathway1')

```

    
    ___Univariate tests results___
    7 tested variables out of 83 variables
    2 DE genes for t=10 and threshold=0.05 (with BH correction)
    Top 5 DE genes (with BH correction): 
    CTSA, CYP51A1





    CTSA       7.042543e-23
    DCP1A      1.521817e-07
    DCTD       9.936203e-06
    CYP51A1    3.675143e-04
    DPP7       6.105009e-02
    DHCR7      3.168313e-01
    CTCF       4.795130e-01
    DHCR24     4.795130e-01
    Name: pathway1_ny_lmrandom_m33_basisw_datacondition_t2_pvalBHc, dtype: float64



Save univariate test results afterwards :
- `save_path`: directory in which to save the result .csv file
- `name` : name of the set of univariate test results to save. 
- `overwrite` : overwrite an eventual existing result file if True, add new results in it if False.  


```python
kt.save_univariate_results(save_path = './',name='pathway1',verbose=1)
```

    - Load univariate test results from 
    	dir: ./
    	file: pathway1_ny_lmrandom_m33_basisw_datacondition_univariate.csv
    - Loaded univariate test results : (8, 25)
    - Saving univariate test results of pathway1 in 
    	dir: ./
    	file:pathway1_ny_lmrandom_m33_basisw_datacondition_univariate.csv ((8, 25))



```python

```


```python

```
